Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-02-03 paper/main.tex
writing results sections for new figs

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/schematic}
\caption{\small \textbf{Methods overview.} We used hand-annotated descriptions of each moment of video to fit a topic model.  Three example video frames and their associated descriptions are displayed (top two rows).  Participants later recalled the video (in the third row, we show example recalls of the same three scenes from participant 13).  We used the topic model (fit to the annotations) to estimate topic vectors for each moment of video and each sentence the participants recalled.  Example topic vectors are displayed in the bottom row (blue: video annotations; green: example participant's recalls).  Three topic dimensions are shown (the highest-weighted topics for each of the three example scenes, respectively).  We also show the ten highest-weighted words for each topic.  Figure~\topics~provides a full list of the top 10 words from each of the discovered topics.}
\label{fig:schematic}
\end{figure}
% Odd name for figure in "Results" section rather than "methods" section -PF

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/eventseg}
\caption{\small \textbf{Modelling naturalistic stimuli and recalls.} All panels: darker colors indicate greater values; range: [0, 1].  \textbf{A.} Topic vectors ($K = 100$) for each of the 1976 video timepoints.  \textbf{B.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel A.  Event boundaries discovered by the HMM are denoted in yellow (30 events detected).  \textbf{C.} Average topic vectors for each of the 30 video events. \textbf{D.} Topic vectors for each of 265 sliding windows of sentences spoken by an example participant while recalling the video.  \textbf{E.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel D. Event boundaries detected by the HMM are denoted in yellow (22 events detected).  For similar plots for all participants see Figure~\corrmats.  \textbf{F.} Average topic vectors for each of the 22 recalled events from the example participant.  \textbf{G.} Correlations between the topic vectors for every pair of video events (Panel C) and recalled events (from the example participant; Panel F).  For similar plots for all participants, see Figure~\matchmats.  \textbf{H.} Average correlations between each pair of video events and recalled events (across all 17 participants).  To create the figure, each recalled event was assigned to the video event with the most correlated topic vector (yellow boxes in panels G and H).  The heat maps in each panel were created using \texttt{Seaborn}~\citep{WaskEtal16}.}
\label{fig:model}
\end{figure}
% I'm all for citing software packages, but where what's the distinction between packages we do cite (seaborn, brainiak, sklearn, hypertools, quail, umap, etc.) and don't cite (wordcloud, nilearn, fastdtw, etc.)?

@@

We can also apply two list-learning-native analyses that describe how participants group items in their recall sequences: temporal clustering and semantic clustering \citep[][see \textit{Methods} for details]{PolyEtal09}.  Temporal clustering refers to the extent to which participants group their recall responses according to encoding position.  Overall, we found that sequentially viewed video events were clustered heavily in participants' recall event sequences (mean: 0.767, SEM: 0.029), and that participants with higher temporal clustering scores tended to perform better according to both \cite{ChenEtal17}'s hand-annotated memory scores (Pearson's $r(15) = 0.62,~p = 0.008$) and our model's estimate (Pearson's $r(15) = 0.54,~p = 0.024$).  Semantic clustering measures the extent to which participants cluster their recall responses according to semantic similarity.  We found that participants tended to recall semantically similar video events together (mean: 0.787, SEM: 0.018), and that semantic clustering score was also related to both hand-annotated  (Pearson's $r(15) = 0.65,~p = 0.004$) and model-derived (Pearson's $r(15) = 0.63,~p = 0.007$) memory performance.

% NOTE: 
% Semantic and temporal clustering are both positively correlated with n_events -- r=0.49; p=0.046 and r=0.61; p=0.009, respectively.  
% - Paxton

@@

Statistical models of memory studies often treat memory recalls as binary (e.g. the item was recalled or not) and independent events.  However, our framework produces a content-based model of individual stimulus and recall events, allowing for direct quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel metrics for quantifying naturalistic memory representations:~\textit{precision} and \textit{distinctiveness}.  We define precision as the average correlation between the topic proportions of each recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.56, p = 0.021$) and the number of recall events estimated by our model (Pearson's $r(15) = 0.85, p < 0.001$). A second novel metric we introduce here is distinctiveness, or how unique the recall description was to each video event. We define distinctiveness as 1 minus the average of all non-matching recall events from the video-recall correlation matrix. We hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that this distinctiveness score is related to our model's estimated number of recalled events (Pearson's $r(15) = 0.49, p = 0.046$) but not to the analogous hand-annotated metric (Pearson's $r(15) = 0.31, p = 0.23$).  In summary, using two novel metrics afforded by our approach, we find that participants whose recalls are both more precise and distinct remember more content.
% ------ NOTE: ------
% rephrase/reframe distinctiveness?
% - Paxton

\begin{figure}[tp]
  \centering
  \includegraphics[width=1\textwidth]{figs/precision_distinctiveness}
  \caption{\small \textbf{Novel content-based metrics of naturalistic memory: precision and distinctiveness.} \textbf{A.} A video-recall correlation matrix for a representative participant (17).  The yellow boxes highlight the maximum correlation in each column.  Precision was computed as the average of the maximum correlation in each column.  On the other hand, distinctiveness was defined as the average of everything except for the maximum correlation in each column. \textbf{B.} The (Pearson's) correlation between precision and hand-annotated memory performance. \textbf{C.} The correlation between precision and the number of events recovered by the model ($k$). \textbf{D.} The correlation between distinctiveness and hand-annotated memory performance. \textbf{E.} The correlation between distinctiveness and the number of events recovered by the model ($k$).}
  \label{fig:precision-distinctiveness}
\end{figure}

@@

>>>>

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/schematic}
\caption{\small \textbf{Methods overview.} We used hand-annotated descriptions of each moment of video to fit a topic model.  Three example video frames and their associated descriptions are displayed (top two rows).  Participants later recalled the video (in the third row, we show example recalls of the same three scenes from participant 13).  We used the topic model (fit to the annotations) to estimate topic vectors for each moment of video and each sentence the participants recalled.  Example topic vectors are displayed in the bottom row (blue: video annotations; green: example participant's recalls).  Three topic dimensions are shown (the highest-weighted topics for each of the three example scenes, respectively).  We also show the ten highest-weighted words for each topic.  Figure~\topics~provides a full list of the top 10 words from each of the discovered topics.}
\label{fig:schematic}
\end{figure}
% Odd name for figure in "Results" section rather than "methods" section 
% - Paxton

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/eventseg}
\caption{\small \textbf{Modelling naturalistic stimuli and recalls.} All panels: darker colors indicate greater values; range: [0, 1].  \textbf{A.} Topic vectors ($K = 100$) for each of the 1976 video timepoints.  \textbf{B.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel A.  Event boundaries discovered by the HMM are denoted in yellow (30 events detected).  \textbf{C.} Average topic vectors for each of the 30 video events. \textbf{D.} Topic vectors for each of 265 sliding windows of sentences spoken by an example participant while recalling the video.  \textbf{E.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel D. Event boundaries detected by the HMM are denoted in yellow (22 events detected).  For similar plots for all participants see Figure~\corrmats.  \textbf{F.} Average topic vectors for each of the 22 recalled events from the example participant.  \textbf{G.} Correlations between the topic vectors for every pair of video events (Panel C) and recalled events (from the example participant; Panel F).  For similar plots for all participants, see Figure~\matchmats.  \textbf{H.} Average correlations between each pair of video events and recalled events (across all 17 participants).  To create the figure, each recalled event was assigned to the video event with the most correlated topic vector (yellow boxes in panels G and H).  The heat maps in each panel were created using \texttt{Seaborn}~\citep{WaskEtal16}.}
\label{fig:model}
\end{figure}
% I'm all for citing software packages, but where what's the distinction between packages we do cite (seaborn, brainiak, sklearn, hypertools, quail, umap, etc.) and don't cite (wordcloud, nilearn, fastdtw, etc.)?
% - Paxton

@@

We can also apply two list-learning-native analyses that describe how participants group items in their recall sequences: temporal clustering and semantic clustering \citep[][see \textit{Methods} for details]{PolyEtal09}.  Temporal clustering refers to the extent to which participants group their recall responses according to encoding position.  Overall, we found that sequentially viewed video events were clustered heavily in participants' recall event sequences (mean: 0.767, SEM: 0.029), and that participants with higher temporal clustering scores tended to perform better according to both \cite{ChenEtal17}'s hand-annotated memory scores (Pearson's $r(15) = 0.62,~p = 0.008$) and our model's estimate (Pearson's $r(15) = 0.54,~p = 0.024$).  Semantic clustering measures the extent to which participants cluster their recall responses according to semantic similarity.  We found that participants tended to recall semantically similar video events together (mean: 0.787, SEM: 0.018), and that semantic clustering score was also related to both hand-annotated  (Pearson's $r(15) = 0.65,~p = 0.004$) and model-derived (Pearson's $r(15) = 0.63,~p = 0.007$) memory performance.
% NOTE: 
% Semantic and temporal clustering are both positively correlated with n_events -- r=0.49; p=0.046 and r=0.61; p=0.009, respectively.  
% - Paxton

@@

Statistical models of memory studies often treat recall success as binary (i.e., an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, the feature-rich content of a naturalistic experiences may later be described with many, highly variable levels of success.  Our framework produces a content-based model of individual stimulus and recall events, allowing for direct quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for quantifying naturalistic memory representations:~\textit{precision} and \textit{distinctiveness}.  We define precision as the ``completeness" of recall-- i.e., how fully the presented content was recapitulated in memory.  Under our framework, we quantify this for a given recall event as the correlation between the topic proportions of the recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which we define as the ``specificity" of recall-- i.e., how unique the description of a given section of content was, compared to descriptions for other sections of content.  We quantify this for each recall event as 1 minus the average correlation between the given recall event and all other recall events not matched to the same video event.  In addition to individual events, one may also use these metrics to describe each participant's overall performance.  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.56, p = 0.021$) and the number of recall events estimated by our model (Pearson's $r(15) = 0.85, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that this distinctiveness score is related to our model's estimated number of recalled events (Pearson's $r(15) = 0.53, p = 0.028$).  While we do not find distinctiveness to be related to hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$), this is not entirely surprising given how the hand-annotated memory scores were computed (see \textit{Methods}).

\begin{figure}[tp]
  \centering
  \includegraphics[width=1\textwidth]{figs/precision_distinctiveness}
  \caption{\small \textbf{Novel content-based metrics of naturalistic memory: precision and distinctiveness.} \textbf{A.} The video-recall correlation matrix for a representative participant (17).  The yellow boxes highlight the maximum correlation in each column.  The example participant's overall precision score was computed as the average across correlation values in the yellow boxes.  Their distinctiveness score was computed as the the average (over recall events) of 1 minus the average correlation between each recall event and all other recall events that do not display a box in the same row.  \textbf{B.} The (Pearson's) correlation between precision and hand-annotated memory performance. \textbf{C.} The correlation between distinctiveness and hand-annotated memory performance. \textbf{D.} The correlation between precision and the number of events recovered by the model ($k$). \textbf{E.} The correlation between distinctiveness and the number of events recovered by the model ($k$).}
  \label{fig:precision-distinctiveness}
\end{figure}

@@

Further intuition for the behaviors captured by these two metrics may be gained by directly examining the content of the video and recalls our framework models.  In Figure \ref{fig:precision-detail}, we contrast two participants' recall descriptions of the same video event: one with a high precision score, the other with a low precision score.  From the HMM-identified event boundaries, we recovered the set of annotations describing the content of an example video event (panel B), and divided them into different color-coded sections for each action or feature described.  We then similarly recovered the set of sentences comprising the corresponding recall event for each of the two example participants.  Because the recall sliding windows overlap heavily, and each recall event spans multiple recall timepoints (i.e., windows), we have stripped any sentences from the beginning and end that describe earlier or later video events.  In other words, the sentences displayed in Fig.~\ref{fig:precision-detail}C are a subset of the full recall event text, spanning sentences between the first and last descriptions of content from the example video event.  We then colored all descriptions of the video event's content by the
